cardinalR: Generating Interesting High-Dimensional Data Structures

### 2025-12-19 paper-cardinalR.Rmd
Set up everything to match the sizes

knitr::include_graphics("figures/cardinalR_workflow.png")
```

@@

tsne_data <- read_rds("data/five_clusts/five_clusts_tsne_perplexity_30.rds")

@@

umap_data <- read_rds("data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds")

@@

phate_data <- read_rds("data/five_clusts/five_clusts_phate_knn_5.rds")

@@

trimap_data <- read_rds("data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")

@@

pacmap_data <- read_rds("data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")

@@

pca_data <- read_rds("data/five_clusts/five_clusts_pca.rds")

@@

<!--scripts/five_clusts/03_gen_rmse.R-->
```{r}
error_five_clust <- read_rds("data/five_clusts/five_clusts_hbe.rds")

@@

# Examine the cluster stats
all_stats <- read_rds("data/five_clusts/cluster_stat.rds")
all_stats |>
  pivot_longer(within.cluster.ss:sindex, names_to = "stat", values_to = "value") |>
  ggplot(aes(x=cl, y=value, colour=method)) +
    geom_line() +
    facet_wrap(~stat, ncol=3, scales="free_y") +
    xlab("Number of clusters") +
    ylab("") +
    scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3')) +
    theme_minimal() +
    theme(panel.border = element_rect(fill = 'transparent'),
          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
          axis.ticks.x = element_line(),
          axis.ticks.y = element_line(),
          legend.position = "bottom",
          axis.text.x = element_text(size = 7),
          axis.text.y = element_text(size = 7),
          axis.title.x = element_text(size = 7),
          axis.title.y = element_text(size = 7),
          plot.margin = margin(0, 0, 0, 0))
```

@@

The source material for this paper is available at [github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR). This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template. These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], `gtools` [@gregory2023], `dplyr` [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].

>>>>

# GitHub raw URL of the figure
fig_url <- "https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/figures/cardinalR_workflow.png"

# Determine figure path depending on output
if (knitr::is_html_output()) {
  # HTML can use the URL directly
  fig_path <- fig_url
} else {
  # PDF needs a local copy
  temp_file <- tempfile(fileext = ".png")
  download.file(fig_url, destfile = temp_file, mode = "wb")
  fig_path <- temp_file
}

knitr::include_graphics(fig_path)
```

@@

tsne_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_tsne_perplexity_30.rds", "rb"))

@@

umap_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_umap_n-neigbors_15_min-dist_0.1.rds", "rb"))

@@

phate_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_phate_knn_5.rds", "rb"))

@@

trimap_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_trimap_n-inliers_12_n-outliers_4_n-random_3.rds", "rb"))

@@

pacmap_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds", "rb"))

@@

pca_data <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_pca.rds", "rb"))

@@

<!--scripts/five_clusts/03_gen_rmse.R-->
```{r}
error_five_clust <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/five_clusts_hbe.rds", "rb"))

@@

# Examine the cluster stats
all_stats <- readRDS(url("https://raw.githubusercontent.com/JayaniLakshika/paper-cardinalR/main/data/five_clusts/cluster_stat.rds", "rb"))
all_stats |>
  pivot_longer(within.cluster.ss:sindex, names_to = "stat", values_to = "value") |>
  ggplot(aes(x=cl, y=value, colour=method)) +
    geom_line() +
    facet_wrap(~stat, ncol=3, scales="free_y") +
    xlab("Number of clusters") +
    ylab("") +
    scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3')) +
    theme_minimal() +
    theme(panel.border = element_rect(fill = 'transparent'),
          plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
          axis.ticks.x = element_line(),
          axis.ticks.y = element_line(),
          legend.position = "bottom",
          axis.text.x = element_text(size = 7),
          axis.text.y = element_text(size = 7),
          axis.title.x = element_text(size = 7),
          axis.title.y = element_text(size = 7),
          plot.margin = margin(0, 0, 0, 0))
```

@@

The source material for this paper, including the full datasets and figures, is available at [github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR). This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template. These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], `gtools` [@gregory2023], `dplyr` [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].
